---
title: "CSC-8015 Data Mining Project"
GroupMembers: "Prashant Barhate, Geeta Yedal, Naimisha Doddi"
date: '2022-08-07'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Mining Project in R

## Sales Prediction for Big Mart

### Load Libraries

```{r loadLibraries}
### Utility function to load packages in case its not available, install it. 
usePackage <- function(p) {
  if (!is.element(p, installed.packages()[,1]))
    install.packages(p, dep = TRUE)
  require(p, character.only = TRUE)
}
## Load required libraries
usePackage("ggplot2")
usePackage("plyr")
usePackage("dplyr")
usePackage("corrplot")
usePackage("gmodels")
```

### Load the data set

```{r loadData}
# Previous Success [1,0]
bigmart <- read.csv("/Users/pbarhate/mastersInCS/projects/dataScience/data/bigmart.csv"
                    , stringsAsFactors = TRUE)
```

### Data Exploration

#### Checking attributes and their data types

```{r dataStructure}
# Check structure of the dataset
str(bigmart)
```

#### Observe Distribution of variables in the data set

```{r dataSummary}
# Check Summary of the data 
summary(bigmart)
```

#### Data Exploration using graphs

##### Sales by Item Type

```{r salesByItemType}
# Sales by item type
ggplot(data = bigmart,aes(x = Item_Type,y = Item_Outlet_Sales)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 70,vjust = 0.5)) +xlab("Item Type") +ylab("Sales")

```

##### Sales by year of establishment

```{r salesByYear}
# Sales by year of establishment
ggplot(data = bigmart,aes(x = Outlet_Establishment_Year,y = Item_Outlet_Sales)) + geom_histogram(stat = "identity")  + xlab("Year Of Establishment") + ylab("Sales") 

```

##### Sales by Outlet Types

```{r salesByOutletTypes}
# Box plot showing sales by outlet type
ggplot(bigmart,aes(x = Outlet_Type,y = Item_Outlet_Sales)) + geom_boxplot() + xlab("Outlet Type") + ylab("Item Sales")
```

##### Checking Distribution of Predictor variables

```{r salesDistribution}
# Bar plot showing distribution of predictor variables
ggplot(bigmart,aes(x = Item_Outlet_Sales)) + geom_histogram(colour = "Black",fill = "cornflowerblue")
```

### Data Cleaning

As seen above there are missing data for Item Weight. Let's clean that data. Assuming that each item identifier actually identifies a unique item,hence a unique weight, let's create a data frame containing the mean weights and standard deviations by item identifier

```{r cleanItemWeight}
bigmart <- ddply(bigmart,~Item_Identifier,transform,Item_Weight = ifelse(is.na(Item_Weight), mean(Item_Weight,na.rm = TRUE),Item_Weight))
```

##### Cleanup missing values in Outlet Size
Replace missing values in outlet size to revalued as "other"

```{r cleanOutletSize}
levels(bigmart$Outlet_Size)
levels(bigmart$Outlet_Size)[1] <- "Other"
```

##### Correction of mismatch level in fat content
Correction of mismatched levels in Fat Content Fat content has similar levels for same fat content. We need to categorize them into Low Fat and Regular. 
```{r fatContent}
bigmart$Item_Fat_Content <- mapvalues(bigmart$Item_Fat_Content,from = c("LF","reg"),to = c("Low Fat","Regular")) 
bigmart$Item_Fat_Content <- mapvalues(bigmart$Item_Fat_Content,from = c("low fat"),to = c("Low Fat"))
```
##### Cleanup Item Visibility
If item exists in store, its visibility cannot be zero. hence clean the 0 visibility to median value
```{r itemVisibility}
bigmart$Item_Visibility[bigmart$Item_Visibility == 0] <- NA 
bigmart <- ddply(bigmart,~Item_Identifier,transform,Item_Visibility =     ifelse(is.na(Item_Visibility),median(Item_Visibility,na.rm=TRUE),Item_Visibility))
```
##### Categorize items in high level categories
On close observation of item_identifier, you can see that items begins with DR, FD or NC. If you check these items under item_type, you will realize that these items are categorized in terms of "Drinks", "Food" and "Non Consumables" respectively. Categorizing items in these high level category will be beneficial hence let's add new attribute to our data frame. 
```{r itemCategory}
## create a new vector and add Drink, Food and Non Consumables categories to it
temp <- substr(bigmart$Item_Identifier,1,2)
temp <- gsub("DR","Drinks",temp) 
temp <- gsub("FD","Food",temp)
temp <- gsub("NC","Non-Consumables",temp)
## Add this vector to the data frame
bigmart$Item_Category <- temp
bigmart$Item_Category <- factor(bigmart$Item_Category,levels = c("Drinks","Food","Non-Consumables"))
```
##### Correcting Fat Content for "Non-Consumable" items

Since we realized that there is a category called "Non-Consumable", having fat content for these items won't be correct. Hence lets replace fat content for "Non-Consumable" items to "Not Applicable".
```{r fatContentForNC}
bigmart$Item_Fat_Content <- as.character(bigmart$Item_Fat_Content)
bigmart$Item_Fat_Content <- ifelse(bigmart$Item_Category == "Non-Consumables", "Not Applicable", bigmart$Item_Fat_Content)
bigmart$Item_Fat_Content <- as.factor(bigmart$Item_Fat_Content)
```
##### Determine age of the outlet
The data in  year of establishment won't be helpful in our model however we can calculate outlet age (that is how long the outlet exists in business) based on establishment year. Since this data is captured in 2013, will calculate age of the outlet as of 2013
```{r outletAge}
## Add new variable as "Outlet_Age" and remove "Outlet_Establishment_Year"
bigmart$Outlet_Age <- 2013 - bigmart$Outlet_Establishment_Year
bigmart <- select(bigmart, -c(Outlet_Establishment_Year))
```
#### Split data into training and testing data sets
```{r splitData}
## Create a randomized vector required to split data
set.seed(12340)
train_sample <- sample(c(TRUE, FALSE), nrow(bigmart), replace=TRUE, prob=c(0.8,0.2))
## Check randomized vector
str(train_sample)
## splitting data in training and testing
train_bigmart  <- bigmart[train_sample, ]
test_bigmart   <- bigmart[!train_sample, ]
```
#### Check Correlation between numeric variables
##### Let see the correlation matrix for numeric variables
```{r correlateMatrix}
# Correlation Matrix for numeric variables
corMatrix <- cor(bigmart[1:nrow(train_bigmart),][sapply(bigmart[1:nrow(train_bigmart),], is.numeric)])
corMatrix
```
##### Visualize the correlations with help of correlation plots
```{r correlatePlots}
# Correlation plots
corrplot::corrplot(corMatrix, method="number", type="upper")
```
Looking at correlations between numerical variables one notices a strong positive correlation  
between Item_MRP and Item_Outlet_Sales and a somewhat weaker negative correlation between 
Item_Visibility and Item_Outlet_Sales.

##### Explore more relations using scatter plots

##### Scatter plot of Item_Outlet_Sales vs Item_MRP
```{r outletSalesNMRP}
# colored according to the Outlet type
ggplot(bigmart[1:nrow(train_bigmart),], aes(Item_MRP, Item_Outlet_Sales)) +
  geom_point(size = 2.5, aes(colour = factor(Outlet_Type))) +
  theme(axis.text.x = element_text(angle = 70, vjust = 0.5, color = "black")) + 
  xlab("Item MRP") + 
  ylab("Item Outlet Sales") +
  ggtitle("Item Sales vs Item MRP")
```
##### Scatter plot of Item_Outlet_Sales vs Item_Visibility
```{r outletSalesNVisibility}
# colored according to the Outlet type
ggplot(bigmart[1:nrow(train_bigmart),], aes(Item_Visibility, Item_Outlet_Sales)) +
  geom_point(size = 2.5, aes(colour = factor(Outlet_Type))) +
  theme(axis.text.x = element_text(angle = 70, vjust = 0.5, color = "black")) + 
  xlab("Item Visibility") + 
  ylab("Item Outlet Sales") +
  ggtitle("Item Sales vs Item Visibility")
```

##### Scatter plot of Item_MRP vs Item_Visibility
```{r mrpNVisibility}
# colored according to the Outlet type
ggplot(bigmart[1:nrow(train_bigmart),], aes(Item_Visibility, Item_MRP)) +
  geom_point(size = 2.5, aes(colour = factor(Outlet_Type))) +
  theme(axis.text.x = element_text(angle = 70, vjust = 0.5, color = "black")) +
  xlab("Item Visibility") +
  ylab("Item MRP") +
  ggtitle("Item MRP vs Item Visibility")
```
###### boxplot of  Sales vs. Item type
```{r salesNItemType}
ggplot(bigmart[1:nrow(train_bigmart),], aes(x = Item_Type, y = Item_Outlet_Sales, fill = Outlet_Type)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 70, vjust = 0.5, color = "black")) + 
  xlab("Item type") + 
  ylab("Sales") + 
  ggtitle("Sales vs Item type")
```

##### bar graph of Item_Outlet_Sales vs Item_Type
```{r barSalesNType}
ggplot(train_bigmart, aes(Item_Type, Item_Outlet_Sales/Item_MRP, fill = Outlet_Type)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  theme(axis.text.x = element_text(angle = 70, vjust = 0.5, color = "black")) + 
  xlab("Item Type") + 
  ylab("Item Outlet Sales") +
  ggtitle("Item Sales vs Item type")
```

##### Linear Regression Model with Important Variables
```{r linearModel1}
## Linear Regression model 
lm1 <- lm(data = train_bigmart, Item_Outlet_Sales ~ Item_Weight + Item_Fat_Content + Item_Visibility + Item_Type + Item_MRP  + Outlet_Size + Outlet_Location_Type + Outlet_Type + Item_Category + Outlet_Age)
summary(lm1)
```
##### Another model
```{r lrModel}
lr_model <- lm(data = train_bigmart, Item_Outlet_Sales ~  Item_MRP  + Outlet_Size + Outlet_Location_Type + Outlet_Type  + Outlet_Age)
## Print Summary of the model
summary(lr_model)
```
##### Plot the model to visualize it
```{r plotModel}
## plot the model
plot(lr_model)
```
##### Predict the sales for test data set
```{r salesPrediction}
sales_predict <- predict(lr_model, newdata = test_bigmart)
summary(sales_predict)
```

